Convex optimisation in communications with cvxpy

### 2017-07-17 article_cvxpy.tex
waterfilling removed

\section{Real world applications}
\subsection{Water filling}
\begin{figure}
\includegraphics[width=1.0\linewidth]{water_filling_plot.pdf}
\caption{\label{fig:waterfilling}Water filling solution for three channels.}
\end{figure}

Convex optimisation can be used to solve the classic water filling problem.  In communications, this is where a total power $P$ must be assigned to $n$ different communication channels, with the objective of maximising the total communication rate.  The communication rate of the the $i$th channel, $r_i$, is given by \cite[p.245]{cvxpybook},

\begin{align}
r_i = \log(x_i+\alpha_i),
\end{align}
where $x_i$ is the power allocated to channel $i$ and $\alpha_i$ represents the floor above the baseline at which power can be added to the channel.  From Fig. \ref{fig:waterfilling}, we can see how this is analogous to filling an uneven basin to a constant level.

%Consider a bounded area of contiguous, equal-area patches $i$ of different elevations $\alpha_i$ (Fig.~\ref{fig:waterfilling}). The area is flooded to a certain level. Find the optimal height $x^{*}_i$ that uses a total amount of water of 1---this describes the famous `water filling' (or `water pouring') problem. It is also common to information theory, formulated by allocating power to communication channels, where each transmits at power $p_i$ and bandwidth $\alpha_i$, with channel rate $\log(\alpha_i+p_i)$\cite[p.245]{cvxpybook}.

The optimisation problem is then \cite{cvxpybook}:
\begin{align}
& \text{minimise} 
& & f_{0}(\textbf{x})=-\sum^{n}_{i=1}\log(\alpha_{i}+x_{i}),\quad\boldsymbol{\alpha} \succ \textbf{0}\nonumber\\
& \text{subject to} 
&& \textbf{x} \succeq \textbf{0} , \nonumber\\& && \sum^{n}_{i=1}x_{i}=\mathds{1}^{T}\textbf{x}=P
\end{align}
where symbols $\succ$, $\succeq$ denote elementwise inequalities between vectors or matrices.

This is a convex problem since $f_0(\textbf{x})$ has positive curvature within its domain, and the inequality and equality constraints are convex.

\subsubsection{Example \texttt{cvxpy} code in Python}
To indicate the brevity of code implementation in \texttt{cvxpy} we show an abridged version below:
\lstinputlisting[language=Python]{waterfilling.py}

\subsubsection{Computational complexity}
\texttt{cvxpy} employs different solvers for each problem and so they have different computational complexities; these were assessed by averaging the runtime over many runs.\\ 
In the case of water filling, for problem sizes up to 1000 channels, the computational time scaled linearly, Fig.~\ref{fig:waterfillingcomputcomplex}.Given that we have $n$ items in this problem and we apply the objective function to each channel independently, a computational complexity of $O(n)$ should be expected.

\begin{figure}
\includegraphics[width=0.9\linewidth]{water_filling_1000runs_1000buckets.pdf}
\caption{\label{fig:waterfillingcomputcomplex}Linear plot of computation time (wall time) of water filling against problem size, for 1000 runs.}
\end{figure}

@@

\subsection{Minimum power for a given SINR}
How can we minimise the total power consumption $P$ of $n$ transmitters each with power $p_i$, yet achieve a minimum SINR $\gamma_0$ for all $m$ receivers? This question is relevant to a telecoms company who want to offer a service at a minimum quality standard. We formulate the problem with a given square path gain matrix $G$, background noise level $\sigma$ and maximum power value that any transmitter can reach $P_{\text{max}}$:
		\begin{align*}
				&\underset{p}{\text{minimise}} \quad &&\sum_j p_j\\
				&\text{subject to} \quad &&p_j \leq P_{\max}\\
				& \quad &&p_j \geq 0\\
				& \quad &&\gamma_i \geq \gamma_0&&
		\end{align*}
The desired signal to receiver $i$ is denoted by $S_i = G_{ik}p_k$, while the interference to $i$ is $I_i = \sum_{j\neq k}G_{ij}p_j$. However DCP does not allow division of the optimisation variable $p_i$ so  
\begin{equation*}
  \gamma_i \geq \gamma_0\Longleftrightarrow  \frac{S_i}{\sigma_i + I_i}\geq \gamma_0, \quad \forall i
\end{equation*}
was rearranged to:
\begin{equation*}
S_i-\gamma_0(\sigma_i + I_i)\geq 0, \quad \forall i
\end{equation*} which is DCP.
\subsubsection{Path gain}
The path gain $G_{ij}$ represents the proportion of power that reaches receiver $i$ from transmitter $j$. Supposing that the desired signal to receiver $i$ is from transmitter $k$, the power received is,
\begin{equation}
p_{ik}^{\text{rec}} = G_{ik}p_k.
\end{equation}
The signal-to-interference-plus-noise ratio (SINR) is the strength of the desired signal $S_i$ relative to the interference power $I_i$ plus the background noise $\sigma_i$ at $i$. Denoting the SINR at $i$ by $\gamma_i$, we have
\begin{equation}
\gamma_i = \frac{G_{ik}p_k}{\sigma_i+\sum_{j\neq k}G_{ij}p_j}
=\frac{S_i}{\sigma_i+I_i}.
\end{equation}

In a physical context, the path gain from transmitter $j$ to receiver $i$ will depend on the distance, $d_{ij}$ between them.  Assuming isotropic propagation from each transmitter, the fraction of power from $j$ that reaches $i$ is given by,

@@

\subsubsection{Example code}
The \texttt{cvxpy} implementation is shown below.
\lstinputlisting[language=Python]{minimisePgivenSINR.py}

>>>>

\section{Real world applications}

@@

\subsection{Minimum power for a given SINR}
How can we minimise the total power consumption $P$ of $n$ transmitters each with power $p_i$, yet achieve a minimum SINR $\gamma_0$ for all $m$ receivers? This question is relevant to a telecoms company who want to offer a service at a minimum quality standard. We formulate the problem with a given square path gain matrix $G$, background noise level $\sigma$ and maximum power value that any transmitter can reach $P_{\text{max}}$:
		\begin{align*}
				&\underset{p}{\text{minimise}} \quad &&\sum_j p_j\\
				&\text{subject to} \quad &&p_j \leq P_{\max}\\
				& \quad &&p_j \geq 0\\
				& \quad &&\gamma_i \geq \gamma_0&&
		\end{align*}
(where symbols $\succ$, $\succeq$ denote elementwise inequalities between vectors or matrices.)

The desired signal to receiver $i$ is denoted by $S_i = G_{ik}p_k$, while the interference to $i$ is $I_i = \sum_{j\neq k}G_{ij}p_j$. However DCP does not allow division of the optimisation variable $p_i$ so  
\begin{equation*}
  \gamma_i \geq \gamma_0\Longleftrightarrow  \frac{S_i}{\sigma_i + I_i}\geq \gamma_0, \quad \forall i
\end{equation*}
was rearranged to:
\begin{equation*}
S_i-\gamma_0(\sigma_i + I_i)\geq 0, \quad \forall i
\end{equation*} which is DCP.
\subsubsection{Path gain}
The path gain $G_{ij}$ represents the proportion of power that reaches receiver $i$ from transmitter $j$. Supposing that the desired signal to receiver $i$ is from transmitter $k$, the power received is,
\begin{equation}
p_{ik}^{\text{rec}} = G_{ik}p_k.
\end{equation}
The signal-to-interference-plus-noise ratio (SINR) is the strength of the desired signal $S_i$ relative to the interference power $I_i$ plus the background noise $\sigma_i$ at $i$. Denoting the SINR at $i$ by $\gamma_i$, we have
\begin{equation}
\gamma_i = \frac{G_{ik}p_k}{\sigma_i+\sum_{j\neq k}G_{ij}p_j}
=\frac{S_i}{\sigma_i+I_i}.
\end{equation}
In a physical context, the path gain from transmitter $j$ to receiver $i$ will depend on the distance, $d_{ij}$ between them.  Assuming isotropic propagation from each transmitter, the fraction of power from $j$ that reaches $i$ is given by,

@@

\subsubsection{Example code}
To indicate the brevity of code implementation in \texttt{cvxpy} we show an abridged version below: \lstinputlisting[language=Python]{minimisePgivenSINR.py}
